---
title: "stcls"
author: "Seungmin Seo" 
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stcls}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Functions in stcls 

### knoxtest

A test to quantify a space and time interaction. The test statistic is

$$K = \sum_{i=1}^N \sum_{j=1}^{i-1} d_{ij} t_{ij} $$

where $N$ is the total number of data, $d_{ij}$ denotes space adjacency that 
equal to 1 if $dist(i,j) < \alpha$, otherwise equal to 0. $t_{ij}$ denotes time 
adjacenct that equal to 1 if $dist(i,j) < \beta$, otherwise equal to 0. 
The Monte Carlo permutation method is used for the significant test of $K$.
by fixing coordinate matrix and changing time vector with permutation.
The biggest demerit of this method is critical space and time distances values 
may be assigned subjectively. To alleviate this problem, two paramters are
added `delta_ratio` and `gamma_ratio`.

1. usage

  > knoxtest(geo, time, delta = NULL, gamma = NULL, delta_ratio, gamma_ratio, B)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data
  
  `delta` is a numeric value of critical space distances values
  
  `gamma` is a numeric value of critical time distances values
  
  `delta_ratio` is a scaler for assiging delta as (1-a)th percentile of space 
  distances where a is delta_ratio
  
  `gamma_ratio` is a scaler for assiging gamma as (1-b)th percentile of time 
  distances where b is gamma_ratio
  
  `B` is a number of permutations for Monte Carlo permutation test.

3. return 

  A list of randomized knox test statistic, observed knox text statistic, 
  p value from Monte Carlo permutation test 

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
res1 = knoxtest(geo,time,delta=50,gamma=30,B=999)
res2 = knoxtest(geo,time,delta_ratio=0.1,gamma_ratio=0.02,B=999)
```



### manteltest

A test to quantify a space and time interaction that focused on the problem 
of selecting the critical distances of Knox’s test. The test statistic is

$$
M = \frac{1}{N^2-N-1} \sum_{i=1}^N \sum_{j=1}^N \frac{\left( d_{ij}^s - \bar{d^s} \right)}{s_s} \frac{\left( d_{ij}^t - \bar{d^t} \right)}{s_t}
$$

where $d_{ij}^s, d_{ij}^t$ are the distance between $i$ and $j$ in space and 
time respectively. $s_s, s_t$ are the standard deviations of data in space 
and time, respectively. Mantel test solves multiple scale problem and 
subjectivity problem of knox test.

1. usage

  > manteltest(geo, time, B)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data
  
  `B` is a number of permutations for Monte Carlo permutation test.

3. return 

  A list of randomized Mantel test statistic, observed Mantel text statistic, 
  p value from Monte Carlo permutation test 

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
res = manteltest(geo,time,B=999)
```


### knn_measure

Calculate $d_ij$ and $t_ij$. $d_ij = 1$ when case $j$ is a $k$ nearest neighbor of 
case $i$ in space, otherwise $d_ij = 0$. $t_ij = 1$ when case $j$ is a 
k-nearest neighbor of case $i$ in time, otherwise $t_ij = 0$.

1. usage

  > knn_measure(geo, time, k, loc = T)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data

  `k` the number of k nearest neightbors
  
  `loc` A logical value. If true, return distance stastistic. Otherwise, 
  return time statistic. (default TRUE)

3. return 

  Matrix $d_ij$ or $t_ij$.

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
knn_d = knn_measure(geo,time,k=5)
knn_t = knn_measure(geo,time,k=5,loc=FALSE)
```


### knn_test_stat

Calculate cumulative test statistic and k specific test statistic of 
spatiotemporal k Nearest Neighbors Test. Each test statistics is 
$$D_k = \sum_{i=1}^N \sum_{j=1}^N d_{ij} t_{ij}$$
$$\Delta D_k = D_k - D{k-1}$$

1. usage

  > knn_test_stat(geo, time, K)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data

  `k` the number of k nearest neightbors
  
3. return 

  A vector of test statistics $D_k, lambda_D_k$.

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
res = knn_test_stat(geo,time,K=5)
```


### knn_test

Spatiotemporal k Nearest Neighbors Test with Monte Carlo permutation test.
$D_k$ depends on $k$ while $\Delta D_k$ is independent because it only contained
specific k nearest neighbors. The null hypothesis is that the distribution of 
events is independent from each other in space and time. The demerit of this 
method is that the $k$ value could result in different test results.

1. usage

  > knn_test(geo, time, B, K)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data

  `B` Number of permutations for Monte Carlo permutation test

  `K` the number of k nearest neightbors
  
3. return 

  A list composed of a B*2 matrix of two test statistics $D_k, lambda_D_k$, 
  bserved test statistics, p-value of cumulative test statistic and 
  p-value of k-specific test statistic.

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
res = knn_test(geo,time,B=999,K=5)
```



### stcls_hyp_plot

Compare the hypothesis testing-based clustering methods of the events spatio-temporal data type by histograms with p-value of Monte Carlo permutation test

1. usage

  > knn_test(geo, time, B, K)

2. parameters 

  `geo` is a dataframe with two columns containing geographic(coordinates) data.
  
  `time` is a vector containing time data

  `delta` is a numeric value of critical space distances values
  
  `gamma` is a numeric value of critical time distances values
  
  `delta_ratio` is a scaler for assiging delta as (1-a)th percentile of space 
  distances where a is delta_ratio
  
  `gamma_ratio` is a scaler for assiging gamma as (1-b)th percentile of time 
  distances where b is gamma_ratio
  
  `B` Number of permutations for Monte Carlo permutation test

  `K` the number of k nearest neightbors

3. return 

  4 histograms of know test, mantel test, knn test with cumulative test 
  statistic, knn test with k-specific test statistic with p-value of Monte 
  Carlo permutation test respectively.

4. example 

```{r}
library(stcls)
geo <- matrix(rnorm(1000 * 2), 1000, 2)
time <- rexp(1000)
stcls_hyp_plot(geo, time, delta=50, gamma=30, B=50,K=5)
stcls_hyp_plot(geo, time, delta_ratio = 0.1, gamma_ratio = 0.02, B=50,K=5)
```

